---
title: "Christian Attitudes - Chi-Squared Test"
subtitle: "Course MATH 325"
author: "Lexi Soelberg"
format:
  html:
    self-contained: true
    page-layout: full
    title-block-banner: true
    toc: true
    toc-depth: 3
    toc-location: body
    number-sections: false
    html-math-method: katex
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: hover
    code-tools:
        source: false
        toggle: true
        caption: See code
execute: 
  warning: false
    
---


```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)
library(mosaic)
library(car)
library(tidyverse)
library(pander)
library(DT)
library(dplyr)
library(readr)
library(haven)

gss2021 <- read_sas("../../data/gss2021.sas7bdat", NULL)
```


<!-- Directions

Come up with a chi squared test analysis of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the chi-squared test analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You can use data from 
    - An R data set
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - Do not violate the "Human Subjects Rules" of the university: http://www.byui.edu/executive-strategy-and-planning/institutional-research/research-and-assessments/research-and-surveys

* You need a sample size large enough to meet the requirements of the chi squared test. Unfortunately, the sample size is completely dependent on the situation you select. A good rule of thumb is 10 observations per "combination-group" in your data.

-->


# Background {.tabset}

The 2021 GSS survey sent to Americans had two different questions: [Have you ever had a 'born again' experience?](https://gssdataexplorer.norc.org/variables/1077/vshow) and [Do you try to convince others to accept Jesus?](https://gssdataexplorer.norc.org/variables/1078/vshow) I found these questions with just two responses each `Yes` or `No` to be quite interesting and I wanted to see how they correlated to each other. 

Is there a statistically significant relationship between individuals who report having had a "born again" experience and those who actively try to convince others to accept Jesus?

-----

As a recently returned missionary, I had always loved bringing along my recent converts --those you could say had a `born again` experience-- to my lessons with other friends. There was always such passion in the testimonies of those recent converts. The same was said with all members that we brought to lessons.

I also feel that this goes along well with the verse in the New Testament: [Matthew 28:19-20](https://www.churchofjesuschrist.org/study/scriptures/nt/matt/28?lang=eng#title_number1)

19 ¶ Go ye therefore, and teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost:

20 Teaching them to observe all things whatsoever I have commanded you: and, lo, I am with you alway, even unto the end of the world. Amen.


------


## Counts & Hypothesis

```{r}

table(gss2021$REBORN) %>% 
  setNames(c("Yes", "No")) %>% 
  pander("Responses for Reborn")

table(gss2021$SAVESOUL) %>% 
  setNames(c("Yes", "No")) %>% 
  pander("Responses for SaveSoul")

table(gss2021$REBORN, gss2021$SAVESOUL) %>%
  `rownames<-`(c("Yes", "No")) %>%
  `colnames<-`(c("Yes", "No")) %>%
  pander("Cross-tabulation of Reborn and SaveSoul")

```

As we can see here the counts are almost equal for the Yes and No groups in the `Reborn` and `SaveSoul` respondents. Just over 1400 people who said Yes to being reborn in Christ and near-equal counts for those who are willing to convince others of Christ. 

In the 3rd table we can see who answered yes to both, yes and no to one, no and yes to another and no to both. 


---------

### Hypothesis

$$
  H_0: \text{Having a Born Again experience is not associated (or independent) with trying to convince others of Jesus Christ}
$$


$$ 
  H_a: \text{Having a Born Again experience is associated with trying to convince others of Jesus Christ}
$$

<center>

$\alpha$ = 0.5

</center>

----------


## Table & Chart

```{r}
Jesustable <- table(gss2021$REBORN, gss2021$SAVESOUL)

rownames(Jesustable) <- c("BORN AGAIN", "NOT BORN AGAIN") #these replace the 1, 2, and 3

colnames(Jesustable) <- c("CONVINCE OF CHRIST", "DON'T CONVINCE CHRIST") #these replace the 1 and 2\

Jesustable %>%
  pander(caption="2021 General Social Survey Results", split.table=Inf)
```


```{r}
barplot(
  t(Jesustable), 
  beside = TRUE, 
  legend.text = TRUE, 
  args.legend = list(
    title = "Religiosity", 
    x = "top", 
    inset = 0.05,   
    cex = 0.8,      
    bty = 'n',      
    horiz = FALSE   
  ), 
  col = c("goldenrod1", "chartreuse4"),
  main = "Born Again on Convincing of Christ",
  ylab = "Counts of Respondants"
)

```


In the graph above we might be tempted to observe a correlation of those who are `Born Again` on their willingness to `Convince of Christ`. Let's run a $X^2$ test statistic to see if the findings are significant enough to reject our $H_0$ hypothesis. 

---------

## $X^2$ Test

We will start by looking at our expected and observed counts. 

```{r}


Jesustest <- chisq.test(Jesustable)

Jesustest$expected %>% pander()

```

To make sure these counts are valid and that we can move on to tunning the test statistic we make sure all counts are above 5, which they are.

We are expecting to see more people spread out, what we are actually seeing is the extremes being represented more. The `Born Again` and `Convince of Christ` and the `Not Born Again` and `Don't Convince of Christ` group are the largest with the other two being the outliers. 

```{r}
Jesustest$observed %>% pander()
```
Now we will run our test to see how significant these observed counts are from the expected. 

```{r}

Jesustest <- chisq.test(Jesustable)
Jesustest %>% pander()

```
The test statistic is decently big and the $p$-value is minuscule. This test gives us enough evidence to reject our $H_0$ and conclude that those who are born again are much more willing to preach of Christ and those who are not born again are not willing to preach of Christ. 

--------


# Interpretation

Because we know we found significance with a $X^2$ test statistic of 1227 and a $p$-value of 9.388e-269, which is incredibly low, we know that Having a born again experience and convincing others of Christ are highly correlated. To see which observed counts are affecting this test statistic we will run a Pearson residuals

```{r}

Jesustest$residuals %>% pander

```

We can see from this chart there were more `Born Again` folks who were also in the `Convince of Christ`. Due to this being the largest number **22.25** we can assume a strong relationship. This is where our members and converts are at. This is our group of Pauls. 1030 strong. 

However not everything is looking good for the `Born Again`-ers because in the `Don't Convince of Christ` group there are some notable deviators. **-17.02** the smallest group at 384 people. 

If not equally surprising we are seeing some `Not Born Again` who are looking to spread some Gospel goodness. **-16.75** in the `Not Born Again` and `Convince of Christ`. We were not expecting them to show up to the church potluck and yet here they are, all 413 of them. 

Finally we see the `Not Born Again` and `Don't Convince of Christ` group. They were our largest group with 2084 members. This is a postivie trending relationship, it's closer to what we expected. It can be hard to want to speak of Christ if you don't know what it's like to be made a new creature in Him. 



