---
title: "Car Selling Price - Honda HR-V"
subtitle: "Course MATH 425"
author: "Lexi Soelberg"
format:
  html:
    self-contained: true
    page-layout: full
    title-block-banner: true
    toc: true
    toc-depth: 3
    toc-location: body
    number-sections: false
    html-math-method: katex
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: hover
    code-tools:
        source: false
        toggle: true
        caption: See code
execute: 
  warning: false
---


```{r, message=FALSE, warning=FALSE}
# Load your libraries
library(car)
library(pander)
library(tidyverse)
library(dplyr)
library(mosaic)
library(ggplot2)
library(plotly)
library(DT)


# boxcox to figure out which transformation i need
# y prime
# remove prime and put what 
# predicted value snap the line back. 
#square the prime again after computed value (3rd equation)
# when including boxcox put it away in details and don't present it
# consider log for exponential car
# pick transformation : like the fit (boxcox) or belief system (human logic)
# 3 equations = prime space, what prime is, 

carselling <- data.frame(
  Mileage = c(66000, 27165, 98241, 25, 7, 87208, 5, 80655, 86740, 101714, 145715, 66912, 15411, 77509, 72764, 19136, 21394, 11255, 22291, 31084, 103162, 69057, 9567, 67592, 32238, 147099, 192216, 176835, 138436, 3978),
  Price = c(12999, 17500, 16818, 31324, 32350, 15990, 32434, 14950, 16495, 17687, 13295, 17800, 25000, 17596, 18288, 25575, 27998, 29998, 22998, 25998, 17955, 19590, 23998, 18998, 20404, 14250, 14977, 10977, 13499, 30998),
  Year = c(2019, 2019, 2017, 2025, 2025, 2017, 2025, 2019, 2016, 2017, 2017, 2018, 2018, 2023, 2018, 2017, 2022, 2023, 2024, 2021, 2022, 2022, 2021, 2020, 2021, 2021, 2020, 2016, 2016, 2024)
)

```


#### Background 

The Honda HR-V was introduced as a compact crossover vehicle in 2016 and therefore finding a vehicle with 200k miles is nearly impossible. This vehicle has existed for less than a decade. The data is limited but interesting. Even the oldest models seem to have upheld their value pretty well over the years. The trim for my car specifically the EX-L has great value. 

The cars were found on [KSL.com](https://cars.ksl.com/search/make/Honda/model/HR-V/yearFrom/2016/yearTo/2026/mileageFrom/0/mileageTo/1000000), [Cars.com](https://www.cars.com/shopping/results/?dealer_id=&include_shippable=true&keyword=&list_price_max=&list_price_min=&makes[]=honda&maximum_distance=all&models[]=honda-hr_v&monthly_payment=&page=10&page_size=50&sort=best_match_desc&stock_type=all&trims[]=honda-hr_v-ex_l_awd&trims[]=honda-hr_v-ex_l_w_navigation&year_max=&year_min=2018&zip=83460), [Autotrader](https://www.autotrader.com/cars-for-sale/all-cars/honda/hr-v/rockland-id?endYear=2022&zip=83271), and the best website for nation-wide searching was [TrueCar](https://www.truecar.com/used-cars-for-sale/listings/honda/hr-v/location-rexburg-id/?searchRadius=5000&trim=ex-l).

I collected 30 cars, from 5 miles to 192k and from $11k to $32k, years ranging from 2016-2025.

My Car is a 2021 Honda HR-V EX-L AWD. Her name is Rexie.

--------------

::: {.panel-tabset}

##### Hide Data



##### Show Data

```{r message=FALSE, warning=FALSE}
carselling <- data.frame(
  Mileage = c(66000, 27165, 98241, 25, 7, 87208, 5, 80655, 86740, 101714, 145715, 66912, 15411, 77509, 72764, 19136, 21394, 11255, 22291, 31084, 103162, 69057, 9567, 67592, 32238, 147099, 192216, 176835, 138436, 3978),
  Price = c(12999, 17500, 16818, 31324, 32350, 15990, 32434, 14950, 16495, 17687, 13295, 17800, 25000, 17596, 18288, 25575, 27998, 29998, 22998, 25998, 17955, 19590, 23998, 18998, 20404, 14250, 14977, 10977, 13499, 30998),
  Year = c(2019, 2019, 2017, 2025, 2025, 2017, 2025, 2019, 2016, 2017, 2017, 2018, 2018, 2023, 2018, 2017, 2022, 2023, 2024, 2021, 2022, 2022, 2021, 2020, 2021, 2021, 2020, 2016, 2016, 2024)
)

datatable(carselling, options=list(lengthMenu = c(5,10,15))) %>%
formatStyle(names(carselling), color="mediumseagreen")
```

-----

:::

#### Hypothesis


This analysis attempts to model the price I can expect to sell my car at based on the mileage using a simple linear regression.

$$
  \underbrace{Y_i}_\text{Predicted Price} = {\beta_0} + {\beta_1} \underbrace{X_i}_\text{Mileage} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2)
$$


The hypotheses for my study consider the slope of the regression model $\beta_1$. If the slope is zero then there is not a meaningful relationship between deep and quality sleep.

$$
H_0: \beta_1 = 0 \\  
H_a: \beta_1 \neq 0
$$

<center> $\alpha$ = 0.05 </center>


# HR-V Linear Regression

::: {.panel-tabset}

## Normal Regression

### Graph 


<center>
```{r message=FALSE, warning=FALSE}

carsell.lm <- lm(Price ~ Mileage, data = carselling)
n <- coef(carsell.lm)

mypred.cc <- predict(carsell.lm, data.frame(Mileage=21000), interval="confidence")
mypred.cp <- predict(carsell.lm, data.frame(Mileage=21000), interval="prediction")


carsellm.log <- lm(log(Price) ~ Mileage, data = carselling)
b.log <- coef(carsellm.log)


carsell.lm <- lm(Price ~ Mileage, data = carselling)
n <- coef(carsell.lm)


mypred.c <- exp(predict(carsellm.log, data.frame(Mileage=120000), interval="confidence"))
mypred.p <- exp(predict(carsellm.log, data.frame(Mileage=120000), interval="prediction"))


carsell.ggplot <- ggplot(carselling, aes(y = Price, x = Mileage, color = Price, label = Year)) +
  geom_point(pch = 16, bg = "white", size = 3) + 
  stat_function(fun = function(x) n[1] + n[2] * x, color = "lightseagreen", size = 1.5) +
  
  stat_function(fun = function(x) exp(b.log[1] + b.log[2] * x), color = "cadetblue1", size = 1.5) + 
  scale_color_gradient(low = "steelblue1", high = "steelblue4") +
  labs(
    title = "Honda HR-V EX-L Prices and Mileages Over The Years<br><sup>2016 - 2025</sup>",
    x = "Mileage", 
    y = "Price ($)"
  ) +
  ylim(5000, 40000) + 
  theme_light() +
  geom_text(
    aes(x = 21000, y = 16000, label = "My Car"),
    color = "forestgreen", nudge_y = -2000, size = 3.75
  ) +
  geom_point(
    aes(x = 21000, y = 16000, label = 2021), 
    color = "forestgreen", size = 3.5
  ) +
  geom_point(
    aes(x = 120000, y = predict(carsell.lm, data.frame(Mileage = 120000, label = 2021))),
    color = "darkgoldenrod2", size = 3.5
  ) +
  geom_text(
    aes(x = 120000, y = predict(carsell.lm, data.frame(Mileage = 120000)), 
        label = "When I Sell My Car"),
    color = "darkgoldenrod3", nudge_y = -2500, nudge_x = -10000,  size = 3.75
  ) +
geom_segment(
  aes(x = 21000, y = 16000, xend = 120000, yend = predict(carsell.lm, data.frame(Mileage = 120000))), 
  color = "goldenrod", lwd = 1
) +
  geom_point() + 
  stat_function(fun=function(x) exp(n[1] + n[2]*x)) + 
  geom_segment(aes(x=120000, xend=120000, y=mypred.c[2], yend=mypred.c[3]),
               alpha=0.1, color="goldenrod2", lwd=3) + 
  geom_segment(aes(x=120000, xend=120000, y=mypred.p[2], yend=mypred.p[3]),
               alpha=0.01, color="firebrick", lwd=2)+
  geom_point() + 
  stat_function(fun=function(x) exp(n[1] + n[2]*x)) + 
  geom_segment(aes(x=21000, xend=21000, y=mypred.cc[2], yend=mypred.cc[3]),
               alpha=0.1, color="forestgreen", lwd=3) + 
  geom_segment(aes(x=21000, xend=21000, y=mypred.cp[2], yend=mypred.cp[3]),
               alpha=0.01, color="steelblue", lwd=2)
  
ggplotly(carsell.ggplot, tooltip = c("y", "x", "label"))

```
</center>


```{r}
CarSellPredict <- predict(carsell.lm, data.frame(Mileage=120000))

pander(paste("Predicted Price When Selling:", round(CarSellPredict, 2)))
```


The green dot represents my car when I bought it ($16k - 21k miles) and the yellow is when I plan on selling it at 120k miles for the predicted price of $15,255.59.

Because I got my car at such an amazing price when I bought it, I can run it for another 100k miles and sell it for almost the same price. Hondas are super reliable so I will do exactly that. 

#### Linear Matheatical Model

$$
  \underbrace{\hat{Y}_i}_\text{Pred Price} = \text{27124} + \text{-0.0989} \underbrace{X_i}_\text{Mileage}
$$


### Linear Summary Results

```{r}
carsell.lm <- lm(Price ~ Mileage, data=carselling)
summary(carsell.lm) %>% pander
```

### Linear Interpretation

Our $B_1$ hypothesis The estimated slope of this regression is significantly different from zero at -9.890e-02 ($p$ = 4.88e-09 < $\alpha$). 
There is a significant negative relationship between a car's price based on its mileage. This is what we were looking for. For our HR-Vs that relationship looks like for every mile the car is driven its value depreciated by $0.0989 (approx 10 cents)

Also because our y-intercept is also significant $p$ = 1.645e-21 < $\alpha$ we can conclude that if a HR-V has 0 miles on it it will cost about $27,120.

The correlation coefficient or $R^2$ shows a fairly moderate tightness of the residuals to $\hat{Y_i}$, $R^2$ = 7.11


### Y-Prime Summary Results



<center> $\lambda$ =0 </center>



```{r}

carsellm.log <- lm(log(Price) ~ Mileage, data = carselling)
summary(carsellm.log) %>% pander

```

#### Y-Prime Mathematical Model

$$
\widehat{Y}_i' = 10.21 + -4.856e-06 X_i
$$
$$
\log(\hat{Y}_i) = 10.21 + -4856e-06 X_i
$$

```{r}
carsellm.log <- lm(log(Price) ~ Mileage, data = carselling)
summary(carsellm.log) %>% pander
```

EXP fix


```{r}

exp(coef(carsellm.log))

```

Our final interpretation of $\hat{Y_i}$ is:

$$
\hat{Y}_i = \text{exp}(2.7112 + -0.099 X_i)
$$


### Log Interpretation

When we snapped log back with exp we find the original linear regression results. 

-------

### Appropriateness of Linear Regression

<center>

```{r fig.height=3}

carsell.lm <- lm(Price ~ Mileage, data=carselling)

par(mfrow=c(1,3))

plot(carsell.lm, which=1)

qqPlot(carsell.lm$residuals, main="Q-Q Plot", col="steelblue3", col.lines="steelblue4", pch= 19, id=FALSE)

plot(carsell.lm$residuals, main = "Residuals vs. Order")

```

</center>

There are 5 associated assumptions to the linear regression. All but #4 can be checked with these above plots. 

**Assumption 1: Linear Relationship Between X and Y**

No significant linearity is shown in the Residuals vs Fitted plot, although 27, 1 and 2 are outliers they do not greatly affect the overal linear regression.  

**Assumption 2: Normal Distribution of Error Terms**

We can see a slight skew in the normality of the data as shown in the Q-Q plot. This does not affect the outcome of this model. 

**Assumption 3: Constant Variance (X values)**

There does not seem to be a pattern in the data, showing constant variance.

**Assumption 4: Fixed X Values**

We can assume the prices and mileages of the cars are accurate and consistent measurements of actual prices of HR-V cars. 

**Assumption 5: Independent Error Terms**

The Residuals vs Order plot shows the chaos in the data. Error terms do not appear to be correlated. There is not a clear pattern in the plot. 


There is not sufficient evidence shown in the above plots to discredit the whole linear regression analysis. 

------

: )

## Appendix

```{r}
boxCox(carsell.lm)
```
I'm going to run a log because of this boxCox assumption


------

: )

###### Credits

Thank you Brother Saunders for helping with the codes on Friday to make the boxCox and different lines. 

:::