---
title: "Split Plot - Marshmallow Shooters"
subtitle: "Course MATH 326"
author: "Lexi Soelberg"
format:
  html:
    self-contained: true
    page-layout: full
    title-block-banner: true
    toc: true
    toc-depth: 3
    toc-location: body
    number-sections: false
    html-math-method: katex
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: hover
    code-tools:
        source: false
        toggle: true
        caption: See code
execute: 
  warning: false
---


```{r, warning=FALSE, message=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)
library(tidyverse)
library(lattice)
library(dplyr)
library(ggplot2)
library(knitr)
library(swirl)

# Record your data from your own mini experiment in Excel.
# Save the data as a .csv file in the Data folder of the Statistics-Notebook.


mm <- read_csv("../../data/MarshmallowSplitPlot.csv")

mm$angle <- as.factor(mm$angle)

mm$person <- as.factor(mm$person)

mm$marshmallow <- as.factor(mm$marshmallow)

```



## Background 

:::{.panel-tabset}

### Introduction & Methods

We randomized our data via the excel and r. We labeled 10 different marshmallows (our blocking factor) and randomized each one to one of our between factor and each marshmallow received our within factor.
Our between blocks factor is Cali vs Stephanie shooting the marshmallows off of a spoon and our within blocks if if they shot the mashmallows sitting or standing. We taped a target to the floor and measured how close the marshmallow landed to the center of the target (measured in centimeters). 

Stephanie got marshmallows 1, 2, 4, 5, & 7. Cali got marshmallows 3, 6, 8, 9, & 10.

We randomized the order of which marshmallow would be shot off first either sitting and standing, ensuring that each marshmallow was shot off both sitting and standing angles. 

marshmallow is nested within person. 

### Data

```{r message=FALSE, warning=FALSE}
library(dplyr)

mm <- mm %>%
  select(person, marshmallow, angle, cm, order)
```

```{r, warning=FALSE, message=FALSE}
datatable(mm, options=list(lengthMenu = c(5,10,20)))
```

## Analysis {.tabset .tabset-fade}

### Hypotheses

EFFECT HYPOTHESES

$$
H_0: \text{Person} (\alpha_\text{i}) = 0 \text{ for all } i
$$

$$
H_a:  \text{Person} (\alpha_\text{i}) \ne 0 \text{ for some } i
$$

$$
H_0: \text{Angle}(\gamma_\text{k}) = 0 \text{ for all } k
$$

$$
H_a: \text{Angle}(\gamma_\text{k}) \ne 0 \text{ for some } k
$$

$$
H_0: \text{Person x Angle}(\alpha\gamma)_\text{ik} = 0 \text{ for all } ik
$$

$$
H_a: \text{Person x Angle}(\alpha\gamma)_\text{ik} \ne 0 \text{ for some } ik
$$

------------------------------------------------------------------------

**Split Plot Mathematical Model**

$$
y_{ijk} = \mu + \alpha_i + \beta_{ij} + \gamma_k + (\alpha \gamma)_{ik} + \epsilon_{ijk}
\tag{1}
$$

**Significance level is**

$$
\alpha = 0.1
$$
(Alpha is set to 0.1 because the data was weird and I expect a lot of variation).
------------------------------------------------------------------------

### ANOVA (Split Plot)

```{r, warning=FALSE, message=FALSE}
mm.aov <- aov(cm ~ person + Error(marshmallow) + angle +
                person:angle, data = mm)
summary(mm.aov) %>% 
  pander()
```

We are seeing that only the angle is significant with a 0.1 alpha 0.09 < $\alpha$. Let's take a closer look with graphs and numeric summaries. 

------------------------------------------------------------------------

### Graphs

**Person Graph**

```{r, warning=FALSE, message=FALSE}
## lets go a ggplot that's real sexy like

library(ggplot2)
library(dplyr)

ggplot(mm, aes(x=person, y=cm)) +
  geom_boxplot(aes(fill = person)) +
  theme_bw() +
  scale_fill_manual(values = c("cadetblue1", "seagreen", "gray")) +
  labs(title = "Distance to Target Achieved by Stephanie and Cali",
       fill = "Person",
       x = "Person",
       y = "Distance from Center of Target (cm)")

```

The averages are very close as depicted in this graph and there are is not too much variation either. 

**Person Numerical Summary**

```{r, warning=FALSE, message=FALSE}
mm %>%
  group_by(person) %>%
  summarise(
    Average = mean(cm, na.rm = TRUE),
    Min = min(cm, na.rm = TRUE), 
    Med = median(cm, na.rm = TRUE), 
    Max = max(cm, na.rm = TRUE),
    StDev = sd(cm, na.rm = TRUE), 
    SampleSize = n()
  ) %>%
  pander()
```
Not significant. 

------------------------------------------------------------------------

**Angle Graph**

```{r, warning=FALSE, message=FALSE}
## lets go a ggplot that's real sexy like

library(ggplot2)
library(dplyr)

ggplot(mm, aes(x=angle, y=cm)) +
  geom_boxplot(aes(fill = angle)) +
  theme_bw() +
  scale_fill_manual(values = c("firebrick", "goldenrod")) +
  labs(title = "Angle Affecting Distance to Target",
       fill = "Angle",
       x = "Angle",
       y = "Distance from Center (cm)")

```

Stand has so much variance, the marshmallow shot off while standing had so much variance from the target whereas sitting led to closer standard deviation as depicted in the numerical summary below.

**Angle Numerical Summary**

```{r, warning=FALSE, message=FALSE}
mm %>%
  group_by(angle) %>%
  summarise(
    Average = mean(cm, na.rm = TRUE),
    Min = min(cm, na.rm = TRUE), 
    Med = median(cm, na.rm = TRUE), 
    Max = max(cm, na.rm = TRUE),
    StDev = sd(cm, na.rm = TRUE), 
    SampleSize = n()
  ) %>%
  pander()
```

These averages look very different, and from the anova test we see that it is significant.

------------------------------------------------------------------------

**Interaction Graph**

```{r, warning=FALSE, message=FALSE}

xyplot(cm ~ angle, data = mm, groups=person, type=c("p","a"), auto.key=TRUE, 
        main="How Angle by Person Affects Accuracy to Target", 
        ylab="Distance from Center (cm)", 
        xlab="Angle")
```

Stand has the closest to the target over sit with Cali having the marshmallow that got the closest to the target. 

**Interaction Numerical Summary**

```{r, message=FALSE, warning=FALSE}
mm %>%
  group_by(person, angle) %>%
  summarise(
    Average = mean(cm, na.rm = TRUE),
    Min = min(cm, na.rm = TRUE), 
    Med = median(cm, na.rm = TRUE), 
    Max = max(cm, na.rm = TRUE),
    StDev = sd(cm, na.rm = TRUE), 
    SampleSize = n()
  ) %>%
  pander()
```
The standard deviation on standing angle is crazy, but the averages appear much lower than sitting angle. 

------------------------------------------------------------------------

### Diagnostics

```{r, warning=FALSE, message=FALSE}

library(car)

par(mfrow=c(1,3))
qqPlot(mm.aov$marshmallow$residuals, main="Block - Normality", id=FALSE)
plot(mm.aov$marshmallow$fitted.values, mm.aov$marshmallow$residuals, main="Block - Constant Variance", pch=16)
plot(mm.aov$marshmallow$residuals, main="Block - Independece", id=FALSE)

```
It looks like for the blocks, nearly every assumption has been violated. Normality has not been met, constant variance shows a megaphone type pattern, and the independence plot doesn't look randomized, there may be trends. 

```{r, warning=FALSE, message=FALSE}
library(car)

par(mfrow=c(1,3))

qqPlot(mm.aov$Within$residuals,  main="Residuals - Normality", id=FALSE)

plot(mm.aov$Within$fitted.values, mm.aov$Within$residuals, main="Residuals - Constant Variance",pch=16)

plot(mm.aov$Within$residuals, main="Residuals - Constant Variance", id=FALSE)

```

These QQ plot shows the normality of the data within the bounds so we can assume that normality assumptions are met.

There is not a megaphone pattern in the residuals vs fitted plot, there is constant variance.

The 3rd plot depicts independence, there are no obvious trends in the data so we can conclude that independence requirements have been met.

We can conclude that all requirements have been met and that we can continue with the interpretation and post hoc tests of the experiment.

------------------------------------------------------------------------

### Post Hoc Tests

#### Grouping Comparions

**Interaction Groupings**

```{r, warning=FALSE, message=FALSE}
library(emmeans)

mm.means <- emmeans(mm.aov, ~ person * angle, infer = c(TRUE, TRUE))

myContrasts <- contrast(
  mm.means,
  list(
    CaliSitvsOthers = c(1, -0.3, -0.3, -0.3),
    CaliStandvsOthers   = c(-0.3, 1, -0.3, -0.3),
    StephanieSitvsOthers = c(-0.3, -0.3, 1, -0.3),
    StephanieStandvsOthers = c(-0.3, -0.3, -0.3, 1)
  ),
  adjust = "scheffe"
)

myContrasts

```

Again like the statistical analysis the anova shows that none are significant. These post hocs are unnecessary, but I wanted to see just in case.

```{r, warning=FALSE, message=FALSE}
library(agricolae)

mm$Group <- with(mm, interaction(person, angle))
mm.aov2 <- aov(cm ~ Group, data = mm)
scheffe.test(mm.aov2, "Group", console = TRUE)
```

This scheffé test shows that each of these groups are not different for each other. 

------------------------------------------------------------------------

:::

## Interpretation

All in all I think this experiment was very silly and it didn't lead to much significant results. This wasn't my favorite or best design and I feel that if we had more blocks, more marshmallows we might have been able to see just more results that ultimately could have been significant. 

------------------------------------------------------------------------

: )
