---
title: "I/O Dataset"
author: "Lexi Soelberg"
format:
  html:
    self-contained: true
    page-layout: full
    title-block-banner: true
    toc: true
    toc-depth: 3
    toc-location: body
    number-sections: false
    html-math-method: katex
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: hover
    code-tools:
        source: false
        toggle: true
        caption: See code
execute: 
  warning: false
    
---

```{r}
pacman::p_load(
  rio,         
  tidyverse,   
  janitor,     
  lubridate,   
  openxlsx,    
  psych, 
  readr, 
  lubridate, 
  pander, 
  car, 
  lsr, 
  plotly
)
```

# Data

::: {.panel-tabset}

```{r}
pre_raw  <- read_csv(
  "../data/io_pre.csv", 
  col_types = cols(.default = "c"),
  na = c("", "NA", "#DIV/0!", "#VALUE!") # <--- Turns Excel errors into blank NAs!
) %>% 
  slice(-c(1, 2)) %>% 
  type_convert()

post_raw <- read_csv(
  "../data/io_post.csv", 
  col_types = cols(.default = "c"),
  na = c("", "NA", "#DIV/0!", "#VALUE!")
) %>% 
  slice(-c(1, 2)) %>% 
  type_convert()

```
```{r}


pre_clean <- pre_raw %>%
  filter(!is.na(Q14), Q14 != "") %>%
  mutate(
    name_key = str_to_lower(str_squish(str_remove_all(Q14, "[[:punct:]]"))),

    StartDate = parse_date_time(StartDate, orders = c("ymd HMS", "mdy HMS", "ymd HM", "mdy HM"))
  ) %>%
  arrange(name_key, StartDate) %>%
  distinct(name_key, .keep_all = TRUE)


post_clean <- post_raw %>%
  filter(!is.na(Q14), Q14 != "") %>%
  mutate(
    name_key = str_to_lower(str_squish(str_remove_all(Q14, "[[:punct:]]"))),

    StartDate = parse_date_time(StartDate, orders = c("ymd HMS", "mdy HMS", "ymd HM", "mdy HM"))
  ) %>%
  arrange(name_key, StartDate) %>%
  distinct(name_key, .keep_all = TRUE)


scale_names <- pre_clean %>% 
  select(starts_with("Pre_")) %>% 
  names() %>% 
  str_remove("^Pre_")

paired_cols <- c(rbind(paste0("Pre_", scale_names), paste0("Post_", scale_names)))


demo_cols <- c("Q14", "Q20_1", "Q22", "Q24", "Q26", "Q28", "Which class?", "Q34", "Q32_1", "Q100_1", "Q104", "Q102")

final_io_data <- inner_join(
  pre_clean, 
  post_clean, 
  by = "name_key", 
  suffix = c("_pre", "_post")
) %>%
  mutate(participant_id = paste0("P", sprintf("%03d", row_number()))) %>%
  
  mutate(across(
    c(starts_with("Pre_"), starts_with("Post_")), 
    ~ na_if(as.numeric(.x), 8)
  )) %>%
  
  # FIX: Only filter on Pre-class, and safely preserve students who left it blank (NA)
  filter(is.na(`Which class?_pre`) | `Which class?_pre` != 1) %>% 
  
  select(
    participant_id,
    any_of(paired_cols), 
    any_of(c(demo_cols, paste0(demo_cols, "_pre"), paste0(demo_cols, "_post")))
  ) %>%
  relocate(Q14_pre, Q14_post, .after = participant_id)

```


```{r}
write_csv(final_io_data, "../data/bwo_check_io_data.csv")
```

:::

## Individual Variables

All the variables we are looking at in this dataset. 

::: {.panel-tabset}

### Human Nature

#### Human Nature Normality

```{r}
hn_diff <- final_io_data$`Pre_Human Nature` - final_io_data$`Post_Human Nature`


qqPlot(hn_diff, main = "Q-Q Plot of Human Nature Differences")
```

#### Human Nature Test

```{r}
t.test(final_io_data$`Pre_Human Nature`, final_io_data$`Post_Human Nature`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Human Nature`, final_io_data$`Post_Human Nature`, method = "paired") 

```

#### Human Nature Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Human Nature", "Post_Human Nature")])
names(long_data) <- c("Human_Nature", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Human Nature", "Post_Human Nature"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Human_Nature,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("steelblue2", "steelblue4")
) %>%
  layout(
    title = "Human Nature Changes Across Semester", 
    yaxis = list(title = "Total Average Human Nature"),
    xaxis = list(title = "Timepoint")
  )

```


### Locus of Control

#### Locus of Control Normality

```{r}

loc_diff <- final_io_data$`Pre_L of C` - final_io_data$`Post_L of C`


qqPlot(loc_diff, main = "Q-Q Plot of Locus of Control Differences", col = "darkgreen", col.lines = "palegreen3")

```

#### Locus of Control Test

```{r}
t.test(final_io_data$`Pre_L of C`, final_io_data$`Post_L of C`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_L of C`, final_io_data$`Post_L of C`, method = "paired") 
```

#### Locus of Control Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_L of C", "Post_L of C")])
names(long_data) <- c("Locus_of_Control", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_L of C", "Post_L of C"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Locus_of_Control,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("palegreen3", "darkgreen")
) %>%
  layout(
    title = "Locus of Control Changes Across Semester", 
    yaxis = list(title = "Total Average Locus of Control"),
    xaxis = list(title = "Timepoint")
  )
```

### Efficacy

#### Efficacy Normality

```{r}
eff_diff <- final_io_data$`Pre_Efficacy` - final_io_data$`Post_Efficacy`

qqPlot(eff_diff, main = "Q-Q Plot of Efficacy Differences", col = "firebrick3", col.lines = "coral1")
```

#### Efficacy Test

```{r}
t.test(final_io_data$`Pre_Efficacy`, final_io_data$`Post_Efficacy`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Efficacy`, final_io_data$`Post_Efficacy`, method = "paired") 
```

#### Efficacy Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Efficacy", "Post_Efficacy")])
names(long_data) <- c("Efficacy", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Efficacy", "Post_Efficacy"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Efficacy,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("coral1", "firebrick3")
) %>%
  layout(
    title = "Efficacy Changes Across Semester", 
    yaxis = list(title = "Total Average Efficacy"),
    xaxis = list(title = "Timepoint")
  )
```

### Challenge

#### Challenge Normality

```{r}
chal_diff <- final_io_data$`Pre_Challenge` - final_io_data$`Post_Challenge`

qqPlot(chal_diff, main = "Q-Q Plot of Challenge Differences", col = "purple4", col.lines = "mediumpurple2")
```

#### Challenge Test

```{r}
t.test(final_io_data$`Pre_Challenge`, final_io_data$`Post_Challenge`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Challenge`, final_io_data$`Post_Challenge`, method = "paired") 
```

#### Challenge Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Challenge", "Post_Challenge")])
names(long_data) <- c("Challenge", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Challenge", "Post_Challenge"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Challenge,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("mediumpurple2", "purple4")
) %>%
  layout(
    title = "Challenge Changes Across Semester", 
    yaxis = list(title = "Total Average Challenge"),
    xaxis = list(title = "Timepoint")
  )
```

### Barriers

#### Barriers Normality

```{r}
bar_diff <- final_io_data$`Pre_Barriers` - final_io_data$`Post_Barriers`

qqPlot(bar_diff, main = "Q-Q Plot of Barriers Differences", col = "darkorange2", col.lines = "goldenrod1")
```

#### Barriers Test

```{r}
t.test(final_io_data$`Pre_Barriers`, final_io_data$`Post_Barriers`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Barriers`, final_io_data$`Post_Barriers`, method = "paired") 
```

#### Barriers Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Barriers", "Post_Barriers")])
names(long_data) <- c("Barriers", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Barriers", "Post_Barriers"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Barriers,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("goldenrod1", "darkorange2")
) %>%
  layout(
    title = "Barriers Changes Across Semester", 
    yaxis = list(title = "Total Average Barriers"),
    xaxis = list(title = "Timepoint")
  )
```

### Social Integration

#### Social Integration Normality

```{r}
si_diff <- final_io_data$`Pre_Social Integration` - final_io_data$`Post_Social Integration`

qqPlot(si_diff, main = "Q-Q Plot of Social Integration Differences", col = "darkslategrey", col.lines = "turquoise3")
```

#### Social Integration Test

```{r}
t.test(final_io_data$`Pre_Social Integration`, final_io_data$`Post_Social Integration`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Social Integration`, final_io_data$`Post_Social Integration`, method = "paired") 
```

#### Social Integration Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Social Integration", "Post_Social Integration")])
names(long_data) <- c("Social_Integration", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Social Integration", "Post_Social Integration"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Social_Integration,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("turquoise3", "darkslategrey")
) %>%
  layout(
    title = "Social Integration Changes Across Semester", 
    yaxis = list(title = "Total Average Social Integration"),
    xaxis = list(title = "Timepoint")
  )
```

### Tech

#### Tech Normality

```{r}
tech_diff <- final_io_data$`Pre_Tech` - final_io_data$`Post_Tech`

qqPlot(tech_diff, main = "Q-Q Plot of Tech Differences", col = "darkorchid4", col.lines = "mediumorchid2")
```

#### Tech Test

```{r}
t.test(final_io_data$`Pre_Tech`, final_io_data$`Post_Tech`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Tech`, final_io_data$`Post_Tech`, method = "paired") 
```

#### Tech Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Tech", "Post_Tech")])
names(long_data) <- c("Tech", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Tech", "Post_Tech"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Tech,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("mediumorchid2", "darkorchid4")
) %>%
  layout(
    title = "Tech Changes Across Semester", 
    yaxis = list(title = "Total Average Tech"),
    xaxis = list(title = "Timepoint")
  )
```

### Screen

#### Screen Normality

```{r}
screen_diff <- final_io_data$`Pre_Screen` - final_io_data$`Post_Screen`

qqPlot(screen_diff, main = "Q-Q Plot of Screen Differences", col = "deepskyblue4", col.lines = "skyblue2")
```

#### Screen Test

```{r}
t.test(final_io_data$`Pre_Screen`, final_io_data$`Post_Screen`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Screen`, final_io_data$`Post_Screen`, method = "paired") 
```

#### Screen Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Screen", "Post_Screen")])
names(long_data) <- c("Screen", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Screen", "Post_Screen"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Screen,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("skyblue2", "deepskyblue4")
) %>%
  layout(
    title = "Screen Changes Across Semester", 
    yaxis = list(title = "Total Average Screen"),
    xaxis = list(title = "Timepoint")
  )
```

### Job Satisfaction

#### Job Satisfaction Normality

```{r}
js_diff <- final_io_data$`Pre_Job Satisfaction` - final_io_data$`Post_Job Satisfaction`

qqPlot(js_diff, main = "Q-Q Plot of Job Satisfaction Differences", col = "darkolivegreen", col.lines = "chartreuse3")
```

#### Job Satisfaction Test

```{r}
t.test(final_io_data$`Pre_Job Satisfaction`, final_io_data$`Post_Job Satisfaction`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Job Satisfaction`, final_io_data$`Post_Job Satisfaction`, method = "paired") 
```

#### Job Satisfaction Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Job Satisfaction", "Post_Job Satisfaction")])
names(long_data) <- c("Job_Satisfaction", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Job Satisfaction", "Post_Job Satisfaction"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Job_Satisfaction,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("chartreuse3", "darkolivegreen")
) %>%
  layout(
    title = "Job Satisfaction Changes Across Semester", 
    yaxis = list(title = "Total Average Job Satisfaction"),
    xaxis = list(title = "Timepoint")
  )
```

### Affective Committment

#### Affective Committment Normality

```{r}
ac_diff <- final_io_data$Pre_Affective - final_io_data$Post_Affective

qqPlot(ac_diff, main = "Q-Q Plot of Affective Committment Differences", col = "darkred", col.lines = "indianred1")
```

#### Affective Committment Test

```{r}
t.test(final_io_data$Pre_Affective, final_io_data$Post_Affective, paired = TRUE) |> 
  pander()

cohensD(final_io_data$Pre_Affective, final_io_data$Post_Affective, method = "paired") 
```

#### Affective Committment Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Affective", "Post_Affective")])
names(long_data) <- c("Affective_Committment", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Affective", "Post_Affective"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Affective_Committment,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("indianred1", "darkred")
) %>%
  layout(
    title = "Affective Committment Changes Across Semester", 
    yaxis = list(title = "Total Average Affective Committment"),
    xaxis = list(title = "Timepoint")
  )
```

### Normative Committment

#### Normative Committment Normality

```{r}
nc_diff <- final_io_data$Pre_Normative - final_io_data$Post_Normative

qqPlot(nc_diff, main = "Q-Q Plot of Normative Committment Differences", col = "violetred4", col.lines = "plum2")
```

#### Normative Committment Test

```{r}
t.test(final_io_data$Pre_Normative, final_io_data$Post_Normative, paired = TRUE) |> 
  pander()

cohensD(final_io_data$Pre_Normative, final_io_data$Post_Normative, method = "paired") 
```

#### Normative Committment Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Normative", "Post_Normative")])
names(long_data) <- c("Normative_Committment", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Normative", "Post_Normative"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Normative_Committment,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("plum2", "violetred4")
) %>%
  layout(
    title = "Normative Committment Changes Across Semester", 
    yaxis = list(title = "Total Average Normative Committment"),
    xaxis = list(title = "Timepoint")
  )
```

### Continuance Committment

#### Continuance Committment Normality

```{r}
cc_diff <- final_io_data$Pre_Continuance - final_io_data$Post_Continuance

qqPlot(cc_diff, main = "Q-Q Plot of Continuance Committment Differences", col = "sienna3", col.lines = "sandybrown")
```

#### Continuance Committment Test

```{r}
t.test(final_io_data$Pre_Continuance, final_io_data$Post_Continuance, paired = TRUE) |> 
  pander()

cohensD(final_io_data$Pre_Continuance, final_io_data$Post_Continuance, method = "paired") 
```

#### Continuance Committment Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Continuance", "Post_Continuance")])
names(long_data) <- c("Continuance_Committment", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Continuance", "Post_Continuance"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Continuance_Committment,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("sandybrown", "sienna3")
) %>%
  layout(
    title = "Continuance Committment Changes Across Semester", 
    yaxis = list(title = "Total Average Continuance Committment"),
    xaxis = list(title = "Timepoint")
  )
```

### Engagement

#### Engagement Normality

```{r}
eng_diff <- final_io_data$`Pre_Engagement` - final_io_data$`Post_Engagement`

qqPlot(eng_diff, main = "Q-Q Plot of Engagement Differences", col = "darkcyan", col.lines = "lightseagreen")
```

#### Engagement Test

```{r}
t.test(final_io_data$`Pre_Engagement`, final_io_data$`Post_Engagement`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Engagement`, final_io_data$`Post_Engagement`, method = "paired") 
```

#### Engagement Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Engagement", "Post_Engagement")])
names(long_data) <- c("Engagement", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Engagement", "Post_Engagement"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Engagement,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("lightseagreen", "darkcyan")
) %>%
  layout(
    title = "Engagement Changes Across Semester", 
    yaxis = list(title = "Total Average Engagement"),
    xaxis = list(title = "Timepoint")
  )
```

### Safety

#### Safety Normality

```{r}
safety_diff <- final_io_data$`Pre_Safety` - final_io_data$`Post_Safety`

qqPlot(safety_diff, main = "Q-Q Plot of Safety Differences", col = "cadetblue4", col.lines = "cadetblue2")
```

#### Safety Test

```{r}
t.test(final_io_data$`Pre_Safety`, final_io_data$`Post_Safety`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Safety`, final_io_data$`Post_Safety`, method = "paired") 
```

#### Safety Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Safety", "Post_Safety")])
names(long_data) <- c("Safety", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Safety", "Post_Safety"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Safety,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("cadetblue2", "cadetblue4")
) %>%
  layout(
    title = "Safety Changes Across Semester", 
    yaxis = list(title = "Total Average Safety"),
    xaxis = list(title = "Timepoint")
  )
```

### CWB

#### CWB Normality

```{r}
cwb_diff <- final_io_data$`Pre_CWB` - final_io_data$`Post_CWB`

qqPlot(cwb_diff, main = "Q-Q Plot of CWB Differences", col = "rosybrown4", col.lines = "rosybrown2")
```

#### CWB Test

```{r}
t.test(final_io_data$`Pre_CWB`, final_io_data$`Post_CWB`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_CWB`, final_io_data$`Post_CWB`, method = "paired") 
```

#### CWB Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_CWB", "Post_CWB")])
names(long_data) <- c("CWB", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_CWB", "Post_CWB"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~CWB,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("rosybrown2", "rosybrown4")
) %>%
  layout(
    title = "CWB Changes Across Semester", 
    yaxis = list(title = "Total Average CWB"),
    xaxis = list(title = "Timepoint")
  )
```

### Job Search

#### Job Search Normality

```{r}
jobsearch_diff <- final_io_data$`Pre_Job Search` - final_io_data$`Post_Job Search`

qqPlot(jobsearch_diff, main = "Q-Q Plot of Job Search Differences", col = "darkgoldenrod3", col.lines = "khaki3")
```

#### Job Search Test

```{r}
t.test(final_io_data$`Pre_Job Search`, final_io_data$`Post_Job Search`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Job Search`, final_io_data$`Post_Job Search`, method = "paired") 
```

#### Job Search Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Job Search", "Post_Job Search")])
names(long_data) <- c("Job_Search", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Job Search", "Post_Job Search"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Job_Search,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("khaki3", "darkgoldenrod3")
) %>%
  layout(
    title = "Job Search Changes Across Semester", 
    yaxis = list(title = "Total Average Job Search"),
    xaxis = list(title = "Timepoint")
  )
```

:::
## Appendix

::: {.panel-tabset}

### Depressive Symptoms

#### Depressive Symptoms Normality

```{r}
cesd_diff <- final_io_data$`Pre_CESD` - final_io_data$`Post_CESD`

qqPlot(na.omit(cesd_diff), main = "Q-Q Plot of CESD Differences", col = "midnightblue", col.lines = "lightslateblue")
```

#### Depressive Symptoms Test

```{r}
t.test(final_io_data$`Pre_CESD`, final_io_data$`Post_CESD`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_CESD`, final_io_data$`Post_CESD`, method = "paired") 
```

#### Depressive Symptoms Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_CESD", "Post_CESD")])
names(long_data) <- c("CESD", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_CESD", "Post_CESD"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~CESD,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("lightslateblue", "midnightblue")
) %>%
  layout(
    title = "CESD Changes Across Semester", 
    yaxis = list(title = "Total Average CESD"),
    xaxis = list(title = "Timepoint")
  )
```


### Happiness

#### Happiness Normality

```{r}
happiness_diff <- final_io_data$Pre_Happiness - final_io_data$Post_Happiness


qqPlot(happiness_diff, main = "Q-Q Plot of Happiness Differences")
```

#### Happiness Test

```{r}
t.test(final_io_data$Pre_Happiness, final_io_data$Post_Happiness, paired = TRUE) |> 
  pander()

cohensD(final_io_data$Pre_Happiness, final_io_data$Post_Happiness, method = "paired") 

```

#### Happiness Graph

```{r}

long_data <- stack(final_io_data[c("Pre_Happiness", "Post_Happiness")])
names(long_data) <- c("Happiness", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Happiness", "Post_Happiness"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Happiness,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("olivedrab3", "olivedrab4")
) %>%
  layout(
    title = "Happiness Changes Across Semester", 
    yaxis = list(title = "Total Average Happiness"),
    xaxis = list(title = "Timepoint")
  )

```


### Differential

#### Differential Normality

```{r}
diff_var_diff <- final_io_data$`Pre_Differential` - final_io_data$`Post_Differential`

qqPlot(na.omit(diff_var_diff), main = "Q-Q Plot of Differential Differences", col = "darkmagenta", col.lines = "plum3")
```

#### Differential Test

```{r}
t.test(final_io_data$`Pre_Differential`, final_io_data$`Post_Differential`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Differential`, final_io_data$`Post_Differential`, method = "paired") 
```

#### Differential Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Differential", "Post_Differential")])
names(long_data) <- c("Differential", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Differential", "Post_Differential"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Differential,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("plum3", "darkmagenta")
) %>%
  layout(
    title = "Differential Changes Across Semester", 
    yaxis = list(title = "Total Average Differential"),
    xaxis = list(title = "Timepoint")
  )
```

### Life Satisfaction

#### SWLS Normality

```{r}
swls_diff <- final_io_data$`Pre_SWLS` - final_io_data$`Post_SWLS`

qqPlot(na.omit(swls_diff), main = "Q-Q Plot of SWLS Differences", col = "seagreen", col.lines = "palegreen")
```

#### SWLS Test

```{r}
t.test(final_io_data$`Pre_SWLS`, final_io_data$`Post_SWLS`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_SWLS`, final_io_data$`Post_SWLS`, method = "paired") 
```

#### SWLS Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_SWLS", "Post_SWLS")])
names(long_data) <- c("SWLS", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_SWLS", "Post_SWLS"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~SWLS,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("palegreen", "seagreen")
) %>%
  layout(
    title = "SWLS Changes Across Semester", 
    yaxis = list(title = "Total Average SWLS"),
    xaxis = list(title = "Timepoint")
  )
```

### Stress

#### Stress Normality

```{r}
stress_diff <- final_io_data$`Pre_Stress` - final_io_data$`Post_Stress`

qqPlot(na.omit(stress_diff), main = "Q-Q Plot of Stress Differences", col = "firebrick4", col.lines = "lightcoral")
```

#### Stress Test

```{r}
t.test(final_io_data$`Pre_Stress`, final_io_data$`Post_Stress`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Stress`, final_io_data$`Post_Stress`, method = "paired") 
```

#### Stress Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Stress", "Post_Stress")])
names(long_data) <- c("Stress", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Stress", "Post_Stress"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Stress,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("lightcoral", "firebrick4")
) %>%
  layout(
    title = "Stress Changes Across Semester", 
    yaxis = list(title = "Total Average Stress"),
    xaxis = list(title = "Timepoint")
  )
```

### Anxiety

#### Anxiety Normality

```{r}
anx_diff <- final_io_data$`Pre_Anxiety` - final_io_data$`Post_Anxiety`

qqPlot(na.omit(anx_diff), main = "Q-Q Plot of Anxiety Differences", col = "darkorange3", col.lines = "moccasin")
```

#### Anxiety Test

```{r}
t.test(final_io_data$`Pre_Anxiety`, final_io_data$`Post_Anxiety`, paired = TRUE) |> 
  pander()

cohensD(final_io_data$`Pre_Anxiety`, final_io_data$`Post_Anxiety`, method = "paired") 
```

#### Anxiety Graph

```{r}
#| warning: false

long_data <- stack(final_io_data[c("Pre_Anxiety", "Post_Anxiety")])
names(long_data) <- c("Anxiety", "Timepoint")

long_data$Timepoint <- factor(
  long_data$Timepoint, 
  levels = c("Pre_Anxiety", "Post_Anxiety"),
  labels = c("Pre", "Post")
)

plot_ly(
  data = long_data, 
  y = ~Anxiety,
  x = ~Timepoint,
  type = "box", 
  color = ~Timepoint,
  colors = c("moccasin", "darkorange3")
) %>%
  layout(
    title = "Anxiety Changes Across Semester", 
    yaxis = list(title = "Total Average Anxiety"),
    xaxis = list(title = "Timepoint")
  )
```

:::

